Trap spaces of Boolean networks are conflict-free siphons of their Petri net encoding

### 2023-01-29 manuscripts/TCS2022/tcs2022.tex
revising conclusion

In this article we explored and proved for the first time the equivalence between (minimal) trap spaces of a general Boolean network and (maximal) conflict-free siphons of its Petri net encoding.
From this equivalence, we proposed a new approach for the computation of minimal trap spaces of Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
We also proposed the four possible methods using MaxSAT, CLP, ILP, and ASP for implementing the new approach.
The proposed methods were evaluated on many real-world models from the literature and randomly generated models.
The experimental results show that \dots
We believe that this opens up the way to a much better analysis of large Boolean networks, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.

@@

% Compare among various possible methods and obtained insights
Second, the experimental results shown in this paper mostly expose the differences caused by the prime-implicant phase of \texttt{PyBoolNet}. There is much more information required to distinctly study the practical differences between our ASP encoding and that of \texttt{PyBoolNet}, and notably their size/efficiency ratio.
Hence, we plan to conduct experiments on more real-world models or maybe random models that can be randomly generated by using BoolNet~\cite{mussel2010boolnet}.
Such experiments should include a time-course of the number of ASP solutions found for proper comparison of the ASP encodings.

% Update: can borrow many advents in the field of Petri nets
In addition, there are possibly other methods for computing maximal conflict-free siphons in Petri nets, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}. Although these approaches do not directly support the maximal conflict-free siphon computation now, we plan to investigate them in the future. They could replace our ASP program if they outperform it.
However, the current method appears to already perform very well even on the biggest models we have considered.

>>>>

In this article we have explored and proved for the first time the equivalence between (minimal) trap spaces of a general Boolean network and (maximal) conflict-free siphons of its Petri net encoding.
We have shown several important applications of this finding to studying properties of trap spaces in Boolean networks. 
As an important practical application of the equivalence, we have proposed a new approach for the computation of minimal trap spaces of Boolean networks, based on the enumeration of maximal conflict-free siphons of Petri nets.
We have also proposed the four possible methods using MaxSAT, CLP, ILP, and ASP for implementing the new approach.
The proposed methods have been evaluated on many real-world models from the literature and randomly generated models.
The experimental results show that the new approach vastly outperforms all the state-of-the-art methods in terms of general Boolean networks and is comparable to the \texttt{mpbn} method even better in average in term of locally-monotonic Boolean networks.
We believe that this opens up the way to a much better analysis of large Boolean networks, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.

@@

% TODO: Compare among various possible methods and obtained insights
Second, the experimental results shown in this paper mostly expose the differences caused by the prime-implicant phase of \texttt{PyBoolNet}. There is much more information required to distinctly study the practical differences between our ASP encoding and that of \texttt{PyBoolNet}, and notably their size/efficiency ratio.
Hence, we plan to conduct experiments on more real-world models or maybe random models that can be randomly generated by using BoolNet~\cite{mussel2010boolnet}.
Such experiments should include a time-course of the number of ASP solutions found for proper comparison of the ASP encodings.

% TODO: can borrow many advents in the field of Petri nets
In addition, there are possibly other methods for computing maximal conflict-free siphons in Petri nets, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}. Although these approaches do not directly support the maximal conflict-free siphon computation now, we plan to investigate them in the future. They could replace our ASP program if they outperform it.
However, the current method appears to already perform very well even on the biggest models we have considered.
